Prefetching SPARQL query cacher

### 2016-05-27 dissertation/introdefhistory.tex
remove final para
Most of Martins grammar comments

The Semantic Web can be described as a machine-readable Web of Data,
defined by a technology stack defined by the World Wide Web
Consortium. While individual technologies can be given precise,
operational definitions, the above definition is naive and can hardly
be used operationally to derive new knowledge. The Semantic Web is a
complex, human artefact that must be understood not only in technical
terms, but also in social, economic and cultural terms. First, the Web
is not only readable, but also writeable. Whether the essential parts
of a data integration problem will be solved by e.g. pervasive
ontology alignment, a linked open vocabularies approach where many
players adopt a fairly consistent set of vocabularies, or by major
players are able to force convergence towards e.g. \texttt{Schema.org}
is an example of the social, economic and cultural mechanisms that
shape the future of the Semantic Web, see
Section~\ref{sec:semproblems} for a more elaborate
discussion. Likewise, it may not be the W3C Semantic Web stack at all
that achieves success, though unlikely at this point, it may be
something in the extension of microformats\footnote{see
  \url{http://microformats.org/}}, microdata\cite{microdata}, or
something else entirely. For a thoughtful criticism of Linked Data
with a proposed alternative, see \cite{darobin1}, but again, it is
likely that social, economic and cultural factors will be as decisive
as technical. One may say that e.g. microformats cannot help achieve
the visions that may be stated for the Semantic Web, but that again
emphasises visions, not operational definitions.

One such vision is the article by Berners-Lee, Hendler and Lassila in
Scientific American 2001\cite{berners2001semantic}, but it does not attempt such an
operational definition. Neither \cite{semwebroadmap} nor \cite{berners2000weaving} contains
such definitions. However, it is clear that an essential
characteristic of the Web, and by extension, the Semantic Web, is its
\emph{universality} and therefore, any actor should be as free as
possible to adapt it to their visions and needs. A pursuit of a
definition of the term ``Semantic Web'' would require a philosophical
treatment which would be beyond the scope of this thesis, and even
with that, may not be a fruitful exercise from a technical
perspective.

@@

When the RDF working drafts were first published, I was first
inundated by the amount of new text to read, but at the same time I
had a growing realisation that getting information out was not solving
the problem The Norwegian Skeptics Society needed to solve: We needed
to push information into closed minds. My initial plan to address that
was to enable a higher degree of targeting of information, and a Web-wide
conversation. To do that, I wanted to create a large thesaurus of
topics of interest, and create annotations of the type ``this article
is a rebuttal of that article'', and an index that browsers could
query, so that when a user viewed an article, they would also get a
critical context. At the same time, Netscape was open-sourced as
Mozilla, and so, I was quite convinced I could contribute the code
needed for this to be accepted.

@@

However, all the above had happened on my spare time, as the topic of
my study was something else entirely: Cosmology, and so the interest
laid dormant for a few years as I finished my Cand. Scient.-degree. 
After that, I got some limited funding to write what today would be
called a Semantic Content Management System. The project was far too
ambitious and largely a failure, but it got me some valuable
experience with Semantic Web technology.

@@

Unfortunately, it wasn't allotted time to enhance and document the
SPARQL Endpoint further, and it didn't see much practical use. 
I was, however, allowed to join several World Wide Web Consortium
groups. First, the Web Content Label Incubator Group, which was tasked
with discussing a replacement for the then archaic Platform for
Internet Content Selection, and resolved to use RDF for this
purpose, and it transitioned into the POWDER Working Group, which
provided the specifications.

@@

While the motivation and the problem statements of this work
undoubtedly have been influenced by my intellectual investment in the
Semantic Web, the papers should show that the investigation has not.

>>>>

The Semantic Web can be described as a machine-readable Web of Data,
defined by a technology stack managed by the World Wide Web
Consortium. While individual technologies can be given precise,
operational definitions, the above definition is naive and can hardly
be used operationally to derive new knowledge. The Semantic Web is a
complex, human artefact that must be understood not only in technical
terms, but also in social, economic and cultural terms. First, the Web
is not only readable, but also writeable. Whether the essential parts
of a data integration problem will be solved by e.g. pervasive
ontology alignment, a linked open vocabularies approach where many
players adopt a fairly consistent set of vocabularies, or by major
players that are able to force convergence towards e.g. \texttt{Schema.org}
is an example of the social, economic and cultural mechanisms that
shape the future of the Semantic Web, see
Section~\ref{sec:semproblems} for a more elaborate
discussion. Likewise, it may not be the W3C Semantic Web stack at all
that achieves success, though unlikely at this point, it may be
something in the extension of microformats\footnote{see
  \url{http://microformats.org/}}, microdata\cite{microdata}, or
something else entirely. For a thoughtful criticism of Linked Data
with a proposed alternative, see \cite{darobin1}, but again, it is
likely that social, economic and cultural factors will be as decisive
as technical. One may say that e.g. microformats cannot help achieve
the visions that may be stated for the Semantic Web, but that again
emphasises visions, not operational definitions.

One such vision is the article by Berners-Lee, Hendler and Lassila in
Scientific American 2001\cite{berners2001semantic}, but it does not attempt such an
operational definition. Neither \cite{semwebroadmap} nor \cite{berners2000weaving} contain
such definitions. However, it is clear that an essential
characteristic of the Web, and by extension, the Semantic Web, is its
\emph{universality} and therefore, any actor should be as free as
possible to adapt it to their visions and needs. A pursuit of a
definition of the term ``Semantic Web'' would require a philosophical
treatment which would be beyond the scope of this thesis, and even
with that, may not be a fruitful exercise from a technical
perspective.

@@

When the RDF working drafts were first published, I was first
overwhelmed by the amount of new text to read, but at the same time I
had a growing realisation that getting information out was not solving
the problem \textit{The Norwegian Skeptics Society} needed to solve: We needed
to push information into closed minds. My initial plan to address that
was to enable a higher degree of targeting of information, and a Web-wide
conversation. To do that, I wanted to create a large thesaurus of
topics of interest, and create annotations of the type ``this article
is a rebuttal of that article'', and an index that browsers could
query, so that when a user viewed an article, they would also get a
critical context. At the same time, Netscape was open-sourced as
Mozilla, and so, I was quite convinced I could contribute the code
needed for this to be accepted.

@@

However, all the above had happened on my spare time, as the topic of
my study was something else entirely: Cosmology, and so the interest
lay dormant for a few years as I finished my Cand. Scient.-degree. 
After that, I got some limited funding to write what today would be
called a Semantic Content Management System. The project was far too
ambitious and largely a failure, but it got me some valuable
experience with Semantic Web technology.

@@

Unfortunately, management did not permit more time to enhance and document the
SPARQL Endpoint further, and it didn't see much practical use. 
I was, however, allowed to join several World Wide Web Consortium
groups. First, the Web Content Label Incubator Group, which was tasked
with discussing a replacement for the then archaic Platform for
Internet Content Selection, and resolved to use RDF for this
purpose, and it transitioned into the POWDER Working Group, which
provided the specifications.

@@

### 2016-05-27 dissertation/motivation.tex
Most of Martins grammar comments

There is a class of diverse, yet structurally similar, problems people
face nearly every day: When we book a flight, we may be interested in
a romantic getaway, yet we are forced to select the airport of
destination first. We often end up buying the same groceries because
we don't know what will taste well together. Last time I bought a car,
the two most important features was that it had 4 wheel drive and that
the trunk was long enough to fit a full-size baby stroller
length-wise. To find a property to build a home, I took into account
the sun conditions, that it was close to public transport that would
transport me to campus within a certain time frame, close to
ski tracks, abundance of uranium in the ground, number of children for
my children to play with, and an affordable price level. And once the
house was being built, I had to choose the toilet seat depending on
the frame in the wall and even a toilet seat that would fit.

Of fundamental importance is that is that even though each of these
problems are rarely encountered, there are so many such problems that
we encounter them frequently. I would like to see a Semantic Web
where the plumber delivers his offer in a manner so that the open
options can be distinguished from the constraints, where a query can
be formulated to find the area I'd like to build my house based on
such extensive criteria, where the very rare comparison between a
collapsed baby stroller and car trunk can be done, where I will be
able to cook better food without being a chef and where I can book my
travel based on what I want to experience rather than make me choose a
destination based on lacking information.

@@

Technically, it might be an easier problem to solve if the data could
be centralized, and that solution has been argued by
\cite{DBLP:conf/semweb/BetzGHS12}. Nevertheless, even if all their
arguments remain valid from a technical viewpoint, the most important
arguments against centralisation are social: If important components
of the Semantic Web is under the control of some central authority, it
is likely that one cannot publish anything, link to anything, or
develop any application without permission. It may create an unhealthy
power structure, which may create commercial or intellectual
hegemonies. Tim Berners-Lee discussed this in \cite{berners2000weaving} and an
interview in Wired Magazine in 2014\footnote{see
  \url{http://www.wired.co.uk/news/archive/2014-02/06/tim-berners-lee-reclaim-the-web}}
This is the key motivation to still design a distributed Semantic Web,
regardless of the technical difficulties that may arise.

>>>>

There is a class of diverse, yet structurally similar, problems people
face nearly every day: When we book a flight, we may be interested in
a romantic getaway, yet we are forced to select the airport of
destination first. We often end up buying the same groceries because
we don't know what will taste well together. Last time I bought a car,
the two most important features was that it had 4 wheel drive and that
the trunk was long enough to fit a full-size baby stroller
length-wise. To find a property to build a home, I took into account
the sun conditions, that it was close to public transport that would
transport me to campus within a certain time frame, close to
ski tracks, low concentration of uranium in the ground, number of children for
my children to play with, and an affordable price level. And once the
house was being built, I had to choose the toilet seat depending on
the frame in the wall and even a toilet seat that would fit.

Of fundamental importance is that even though each of these
problems are rarely encountered, there are so many such problems that
we encounter them frequently. I would like to see a Semantic Web
where the plumber delivers his offer in a manner so that the open
options can be distinguished from the constraints, where a query can
be formulated to find the area I'd like to build my house based on
such extensive criteria, where the very rare comparison between a
collapsed baby stroller and car trunk can be done, where I will be
able to cook better food without being a chef and where I can book my
travel based on what I want to experience rather than make me choose a
destination based on lacking information.

@@

Technically, it might be an easier problem to solve if the data could
be centralized, and that solution has been argued for by
\cite{DBLP:conf/semweb/BetzGHS12}. Nevertheless, even if all their
arguments remain valid from a technical viewpoint, the most important
arguments against centralisation are social: If important components
of the Semantic Web are under the control of some central authority, it
is likely that one cannot publish anything, link to anything, or
develop any application without permission. It may create an unhealthy
power structure, which may create commercial or intellectual
hegemonies. Tim Berners-Lee discussed this in \cite{berners2000weaving} and an
interview in Wired Magazine in 2014\footnote{see
  \url{http://www.wired.co.uk/news/archive/2014-02/06/tim-berners-lee-reclaim-the-web}}
This is the key motivation to still design a distributed Semantic Web,
regardless of the technical difficulties that may arise.

### 2016-05-27 dissertation/papersincontext.tex
Most of Martins grammar comments

%TODO: not the Dagstuhl?

This paper is relevant to the topic of the dissertation for two
reasons: First, it shows how hypermedia may be used to drive a
read-write application, not just read-only, which is the focus of most
of the discussion. Second, it provides some of the basis for
\cite{ldf1}, which is important in the query planner.

@@

However, this study merely scratches the surface. To establish a
Design of Experiment-based methodology would be a multi-year project
in its own right, clearly beyond the scope of this
dissertation. Unfortunately, this also implies that I do not have the
rigorous evaluation that I would preferred to have in this
dissertation, it is unattainable.

@@

This work was the outcome of several hackathons that I had organised
in the Perl community. It is not very well known that Gregory Todd
Williams' SPARQL implementation written in Perl is one of the SPARQL
1.1 reference implementations, and is known to be fully compliant with
the specification. Through our work, and the availability of
traits-based programming in Perl, see \cite{traits}, we had extensive
discussions on how to leave optimisations to lower levels, e.g. the
underlying database engine, when they had the potential to provide a
better plan or cost estimate. While it is certainly possible to do
this without this new programming paradigm, traits made the
methodology much more comprehensible, and so helped the subsequent
work substantially.

@@

I committed to do the survey after Jürgen Umbrich challenged me to see
if the standards were actually followed. Rather than just survey the
usage for SPARQL Endpoints, I decided to do it for the breadth of the
Semantic Web. This resulted in a research track paper on the Extended
Semantic Web Conference 2015 entitled ``A Survey of HTTP Caching
Implementations on the Open Semantic Web'' with reference
\cite{kjernsmo_survey_2015} and a technical report
\cite{kjernsmo_add_survey_2015} that elaborates on data reduction
methods and statistical methods.

@@

Dissatisfaction with the methodology used in the literature,
specifically benchmarking, to test assertions of performance lead to
exploration of statistical methods in referenced in
Section~\ref{sec:condoe}, and to organise the Empirical 2014 Workshop
at the Extended Semantic Web Conference (which I was unfortunately
prevented from attending myself). Still, frustrated by the slow
progress of the last 16 years, and difficulty with comparing methods
to the more mature natural sciences, I still felt the need to express
a philosophical view of science, which I did by submitting an article
to the ``Negative or inconclusive results in Semantic Web'' at the
Extended Semantic Web Conference 2015. The essay, see
\cite{kjernsmo_noise_2015} was tangential to the Call for Papers, but
the chairs decided to accept it as an interview. With Ruben Verborgh
in the role of the interviewer, we reformulated it as
\cite{kjernsmo_noise_2015_interview}. It was performed with  Jacco van
Ossenbruggen as the interviewer in front of the audience.

>>>>

%TODO: not the Dagstuhl?
% TODO: relate to problems
This paper is relevant to the topic of the dissertation for two
reasons: First, it shows how hypermedia may be used to drive a
read-write application, not just read-only, which is the focus of most
of the discussion. Second, it provides some of the basis for
\cite{ldf1}, which is important in the query planner.

@@

However, this study merely scratches the surface. To establish a
Design of Experiment-based methodology would be a multi-year project
in its own right, clearly beyond the scope of this
dissertation. Unfortunately, this also implies that I do not have the
rigorous evaluation that I would have preferred to have in this
dissertation, it is unattainable.

@@

This work was the outcome of several hackathons that I had organised
in the Perl community. It is not very well known that Gregory Todd
Williams' SPARQL implementation written in Perl is one of the SPARQL
1.1 reference implementations. It is also fully compliant with
the specification. Through our work, and the availability of
traits-based programming in Perl, see \cite{traits}, we had extensive
discussions on how to leave optimisations to lower levels, e.g. the
underlying database engine, when they had the potential to provide a
better plan or cost estimate. While it is certainly possible to do
this without this new programming paradigm, traits made the
methodology much more comprehensible, and so helped the subsequent
work substantially.

@@

I committed to do this survey after Jürgen Umbrich challenged me to see
if the standards were actually followed. Rather than just survey the
usage for SPARQL Endpoints, I decided to do it for the breadth of the
Semantic Web. This resulted in a research track paper on the Extended
Semantic Web Conference 2015 entitled ``A Survey of HTTP Caching
Implementations on the Open Semantic Web'' with reference
\cite{kjernsmo_survey_2015} and a technical report
\cite{kjernsmo_add_survey_2015} that elaborates on data reduction
methods and statistical methods.

@@

Dissatisfaction with the methodology used in the literature,
specifically benchmarking, to test assertions of performance lead to
exploration of statistical methods referenced in
Section~\ref{sec:condoe}, and to organise the Empirical 2014 Workshop
at the Extended Semantic Web Conference (which I was unfortunately
prevented from attending myself). Still, frustrated by the slow
progress of the last 16 years, and difficulty with comparing methods
to the more mature natural sciences, I still felt the need to express
a philosophical view of science, which I did by submitting an article
to the ``Negative or inconclusive results in Semantic Web'' at the
Extended Semantic Web Conference 2015. The essay, see
\cite{kjernsmo_noise_2015} was tangential to the Call for Papers, but
the chairs decided to accept it as an interview. With Ruben Verborgh
in the role of the interviewer, we reformulated it as
\cite{kjernsmo_noise_2015_interview}. It was performed with  Jacco van
Ossenbruggen as the interviewer in front of the audience.

### 2016-05-27 dissertation/problems.tex
Most of Martins grammar comments

In \cite{berners2000weaving}, Tim Berners-Lee take schema
heterogeneity for granted and that it would be resolved by using
inference, saying:

@@

\begin{problem}\label{prob:sayanything}
With a centralised ontology approach, how are anybody enabled to say
anything?
\end{problem}

@@

To understand whether a result from a study on e.g. performance is
valid, an empirical study must be conducted, as a SPARQL system is
generally too complex to be evaluated solely by formal
methods. Conventionally, one or more benchmarks have been used for
this purpose, and there exists a number such
benchmarks. Unfortunately, due to this complexity, they have to test a
very large number of parameters, but are unable to do that, and they
may also oversimplify the test by attempting to neutralise certain
optimisation techniques. Further, they have no structured approach to
investigate flaws in the benchmark itself, and have no meaningful
summary statistics that can be used to understand the overall
performance. While one may standardise benchmarks, doing so makes it
impossible to test assertions that are outside of that standard.

@@

It is clear that neither the vision presented in
Section~\ref{sec:motivation} nor the more advanced vision presented in
\cite{berners2001semantic} exists as of today. In both cases, Semantic
Web would have to be deployed at a very different scale than what it
is today. If the Semantic Web is sufficiently successful to realise
these visions, it is also likely that it will be very different from
today in terms of data profiles, workloads, etc. Since data profiles
and workloads are important to device a test to evaluate assertions
about performance, the fundamental problem arises that we currently
test against something that does not exist, and this differs markedly
from the methodology of natural science.

@@

\begin{problem}\label{prob:lawsfuture}
Establish laws of the Semantic Web information universe can reasonably be
expected to have validity far beyond the present.
\end{problem}

@@

My next concern was the implications of the
problems~\ref{prob:zipf}~and~\ref{prob:uniform} in a federated SPARQL
engine. While studying the evaluations done in the relevant literature
in the field, I decided to start an effort to address
Problem~\ref{prob:badstats} first by adopting methodology from the
statistical field of ``Design of Experiments'' (DoE). See
Section~\ref{sec:condoe} for details on this work.

@@

As noted in Section~\ref{sec:motivation} it is clear that evaluating
arbitrary SPARQL queries are in general unsustainable, and it is
likely that the problems in Section~\ref{sec:perfproblems} is partly
to blame for it. Our work, and most of the related work in
Section~\ref{sec:related} approach this work from the direction of
moving stress from the server, but it is also clear that another way
to address Problem~\ref{prob:sparqlcomplex} is for the SPARQL engine
to identify queries it would be too heavy to answer already when
parsing or planning, and also expose that in the service description
so that clients can avoid asking them. The HTTP/2 specification
\cite{rfc7540} has some generic facilities that may be helpful:

@@

I have not approached any of the problems of
Section~\ref{sec:semproblems}, mainly because I think it is an open
question which of these approaches will be adopted in practice, and
whatever the solution may be, their designers should make sure they
are orthogonal to the other problems discussed here. Thus, it was
natural to define them out of scope. 
% TODO: However, I shall assume that no inference is needed in the
% following examples.

>>>>

In \cite{berners2000weaving}, Tim Berners-Lee et al take schema
heterogeneity for granted and postulate that it would be resolved by using
inference, saying:

@@

\begin{problem}\label{prob:sayanything}
With a centralised ontology approach, how can we ensure that anybody is enabled to say
anything?
\end{problem}

@@

To understand whether a result from a study on e.g. performance is
valid, an empirical study must be conducted, as a SPARQL system is
generally too complex to be evaluated solely by formal
methods. Conventionally, one or more benchmarks have been used for
this purpose, and there exists a number of such
benchmarks. Unfortunately, due to this complexity, they have to test a
very large number of parameters, but are unable to do that, and they
may also oversimplify the test by attempting to neutralise certain
optimisation techniques. Further, they have no structured approach to
investigate flaws in the benchmark itself, and have no meaningful
summary statistics that can be used to understand the overall
performance. While one may standardise benchmarks, doing so makes it
impossible to test assertions that are outside of that standard.

@@

It is clear that neither the vision presented in
Section~\ref{sec:motivation} nor the more advanced vision presented in
\cite{berners2001semantic} exists as of today. In both cases, Semantic
Web would have to be deployed at a very different scale than what it
is today. If the Semantic Web is sufficiently successful to realise
these visions, it is also likely that it will be very different from
today in terms of data profiles, workloads, etc. Since data profiles
and workloads are important to devise a test to evaluate assertions
about performance, the fundamental problem arises that we currently
test against something that does not exist, and this differs markedly
from the methodology of natural science.

@@

\begin{problem}\label{prob:lawsfuture}
Establish laws of the Semantic Web information universe that can reasonably be
expected to have validity far beyond the present.
\end{problem}

@@

My next concern were the implications of the
problems~\ref{prob:zipf}~and~\ref{prob:uniform} in a federated SPARQL
engine. While studying the evaluations done in the relevant literature
in the field, I decided to start an effort to address
Problem~\ref{prob:badstats} first by adopting methodology from the
statistical field of ``Design of Experiments'' (DoE). See
Section~\ref{sec:condoe} for details on this work.

@@

As noted in Section~\ref{sec:motivation} it is clear that evaluating
arbitrary SPARQL queries are in general unsustainable, and it is
likely that the problems in Section~\ref{sec:perfproblems} are partly
to blame for it. Our work, and most of the related work in
Section~\ref{sec:related} approach this work from the direction of
moving stress from the server, but it is also clear that another way
to address Problem~\ref{prob:sparqlcomplex} is for the SPARQL engine
to identify queries it would be too heavy to answer already when
parsing or planning, and also expose that in the service description
so that clients can avoid asking them. The HTTP/2 specification
\cite{rfc7540} has some generic facilities that may be helpful:

@@

I have not approached any of the problems of
Section~\ref{sec:semproblems}, mainly because I think it is an open
question which of these approaches will be adopted in practice, and
whatever the solution may be, their designers should make sure they
are orthogonal to the other problems discussed here. Thus, it was
natural to consider them out of scope. 
% TODO: However, I shall assume that no inference is needed in the
% following examples.

### 2016-05-27 dissertation/related.tex
Most of Martins grammar comments

Verborgh et al. \cite{ldf1} share many of the observations and goals
that motivates this study. They introduce a hypermedia system that can
be used to answer individual triple patterns (known as Triple Pattern
Fragments (TPF)). Their stated goal is to transfer as much as possible
of the burden of evaluating a query from the server to the
client. This can happen because a simple pattern match for a single
triple pattern need not have a SPARQL parser, nor a query
planner. Moreover, the server may materialise responses to frequent
triple patterns and store them in a file system, which can be managed
by a simple Web server. The results may also be paged, so that each
individual response may be kept small. However, TPF mandates that
every response must contain a cardinality estimate for every triple
pattern, an operation that may be quite expensive for the server, and
%TODO: ref characteristic sets?
indeed, this information may be so expensive to provide that SPARQL
planners may not provide it to keep the cost of planning itself
down. In \cite{verborgh2014querying}, they demonstrated SPARQL evaluation over TPF,
and further improvements have been made in more recent papers.

Olaf Hartig has a large number of papers that are adjacent to the
present work, including his Ph.D. dissertation \cite{hartig2014querying} where he
explores querying Linked Data, i.e. query data that is not contained
in an a-priory defined collection of RDF data, but where a resources
are traversed to compute a result. Hartig provides solid formal
foundations for such query processing, as well as computational
feasibility, soundness and completeness. The most relevant paper to
this work is \cite{hartig2011caching}.

Acosta et al. \cite{acosta2014shepherd} has ongoing work named ``SHEPHERD'' to
create a SPARQL processor that takes into account observations done by
the SPARQLES survey \cite{buil2013sparql}, to decompose a SPARQL query into
subqueries that is indicated to be easier for the server to evaluate,
and so is well aligned with our goal to ease the load of the
server. It is not quite clear from the brief description in 
\cite{acosta2014shepherd} whether the SHEPHERD system can cache results locally.

Recently, Papailiou et al. \cite{papailiou2015graph} made an extensive
study of SPARQL caching, addressing some of the same problems as the
present work, except for our emphasis on deployment on proxies in the
Internet and integration with HTTP-based caching.  Montoya et
al. \cite{DBLP:journals/corr/MontoyaSMV15} proposes a query federation
system that takes advantage of client-side data replication. Both
these studies claim to support the entire SPARQL query language, but
in both cases, it is not clear that they actually do that.

Umbrich et al. \cite{umbrich2012hybrid} considered a situation where
parts of the graph changes at different paces. For the slow-changing
parts, they prefetch an entire dataset to a local store to relieve
remote endpoints from some processing, and thereby speed up query
responses.

>>>>

Verborgh et al. \cite{ldf1} share many of the observations and goals
that motivate this study. They introduce a hypermedia system that can
be used to answer individual triple patterns (known as Triple Pattern
Fragments (TPF)). Their stated goal is to transfer as much as possible
of the burden of evaluating a query from the server to the
client. This can happen because a simple pattern match for a single
triple pattern need not have a SPARQL parser, nor a query
planner. Moreover, the server may materialise responses to frequent
triple patterns and store them in a file system, which can be managed
by a simple Web server. The results may also be paged, so that each
individual response may be kept small. However, TPF mandates that
every response must contain a cardinality estimate for every triple
pattern, an operation that may be quite expensive for the server, and
%TODO: ref characteristic sets?
indeed, this information may be so expensive to provide that SPARQL
planners may not provide it to keep the cost of planning itself
down. In \cite{verborgh2014querying}, they demonstrated SPARQL evaluation over TPF,
and further improvements have been made in more recent papers.

Olaf Hartig has a large number of papers that are adjacent to the
present work, including his Ph.D. dissertation \cite{hartig2014querying} where he
explores querying Linked Data, i.e. query data that is not contained
in an a-priory defined collection of RDF data, but where resources
are traversed to compute a result. Hartig provides solid formal
foundations for such query processing, as well as computational
feasibility, soundness and completeness. The most relevant paper to
this work is \cite{hartig2011caching}.

Acosta et al. \cite{acosta2014shepherd} have ongoing work named ``SHEPHERD'' to
create a SPARQL processor that takes into account observations done by
the SPARQLES survey \cite{buil2013sparql}, to decompose a SPARQL query into
subqueries that is indicated to be easier for the server to evaluate,
and so is well aligned with our goal to ease the load of the
server. It is not quite clear from the brief description in 
\cite{acosta2014shepherd} whether the SHEPHERD system can cache results locally.

Recently, Papailiou et al. \cite{papailiou2015graph} made an extensive
study of SPARQL caching, addressing some of the same problems as the
present work, except for our emphasis on deployment on proxies in the
Internet and integration with HTTP-based caching.  Montoya et
al. \cite{DBLP:journals/corr/MontoyaSMV15} propose a query federation
system that takes advantage of client-side data replication. Both
these studies claim to support the entire SPARQL query language, but
in both cases, it is not clear that they actually do that.

Umbrich et al. \cite{umbrich2012hybrid} considered a situation where
parts of the graph change at different paces. For the slow-changing
parts, they prefetch an entire dataset to a local store to relieve
remote endpoints from some processing, and thereby speed up query
responses.
